
## Studies stopped because of safety

### Safety by Therapeutic area

```{r, safety_byTA, cache = TRUE}
p_safety_by_ta <- all_tests %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    # filter(predictionType %in% c("reason", "clinical", "stopStatus")) %>%
    filter(comparisonType == "ta") %>%
    mutate(comparison = fct_reorder(comparison, estimate, median)) %>%
    ggplot(aes(x = comparison, y = estimate)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
    width = 0) +
    geom_point() +
    scale_y_log10(name = "Odds Ratio (log)") +
    coord_flip() +
    facet_grid(. ~ prediction) +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.title.y = element_blank()
    )
p_safety_by_ta
```

```{r, safety_main, cache = TRUE}
stars_pval <- function(p_value) {
  unclass(
    symnum(
      p_value,
      corr = FALSE,
      na = FALSE,
      cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05,  1),
      symbols = c("****", "***", "**", "*",  "ns")
    )
  )
}

toplot_safety_main <- all_tests %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    filter(comparisonType %in% c("ta",
    "geneticConstrain", "rnaSpecificity", "lof_tolerance", "interactions")) %>%
    mutate(comparison = fct_reorder(comparison, estimate)) %>%
    mutate(comparison = fct_relevel(
        comparison,
        "Tissue enriched",
        "Group enriched",
        "Tissue enhanced",
        "Low tissue specificity"
    )) %>%
    mutate(comparison = fct_relevel(
        comparison,
        "greaterThan20",
        after = Inf)) %>%
    mutate(comparison = fct_recode(comparison,
        "Very high" = "0",
        "High" = "1",
        "Medium" = "2",
        "Low" = "3",
        "Very low" = "4",
        "Lowest" = "5"
    )) %>%
    mutate(comparison = fct_recode(comparison,
        ">20 partners" = "greaterThan20",
        "11-20 partners" = "from11to20",
        "1-10 partners" = "from1to10",
        "None reported" = "none"
    )) %>%
    mutate(comparison = fct_recode(comparison,
        "Intolerant (pLI > 0.9)" = "LoF intolerant",
        "Tolerant (pLI < 0.1)" = "LoF tolerant"
    )) %>%
    mutate(comparisonType = fct_recode(
        comparisonType,
        "LoF" = "lof_tolerance",
        "Interacting proteins" = "interactions",
        "RNA specificity" = "rnaSpecificity",
        "Indication" = "ta",
        "Genetic Constrain (LOEUF bin)" = "geneticConstrain"
    )) %>%
    mutate(comparisonType = fct_rev(
        fct_reorder(comparisonType, estimate, max, na.rm = TRUE))) %>%
    mutate(comparisonType = fct_relevel(
        comparisonType, "Interacting proteins", after = Inf)) %>%
    mutate(oddsinterval = paste(round(estimate, digits = 3),
    " (",
    round(conf.low, digits = 2),
    "-",
    round(conf.high, digits = 2),
    ")",
    sep = "")) %>%
    mutate(significance = stars_pval(p.value))

p_safety_main_dots <- toplot_safety_main %>%
    ggplot(aes(x = comparison, y = estimate)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
    width = 0, color = "steelblue", size = 1) +
    geom_point(color = "steelblue", size = 3) +
    scale_y_log10(name = "Odds Ratio (log)", position = "right") +
    scale_x_discrete(labels = function(x)
    stringr::str_wrap(x, width = 25)) +
    coord_flip() +
    facet_grid(comparisonType ~ .,
        scale = "free_y",
        space = "free_y",
        switch = "y"
    ) +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.title.y = element_blank(),
        # strip.text.y = element_blank(),
        axis.line = element_blank(),
        # panel.spacing = unit(0, "lines"),
        panel.spacing = unit(-0.1, "lines"),
        strip.placement = "outside",
        strip.background = element_blank(),
        strip.switch.pad.grid = unit(20, "pt")
    )

p_safety_main_numbers <- toplot_safety_main %>%
    mutate(a = formatC(a, format = "d", big.mark = ",")) %>%
    select(comparison,
    comparisonType,
    prediction,
    a,
    oddsinterval,
    significance) %>%
    gather(key, value, -comparison, -comparisonType, -prediction) %>%
    mutate(key = fct_recode(key,
    "n" = "a",
    "OR (95% CI)" = "oddsinterval",
    "Significance" = "significance")) %>%
    ggplot(aes(x = comparison, y = key)) +
    # geom_tile(fill = NA, color = "black") +
    geom_tile(fill = NA) +
    geom_text(aes(label = value)) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 25)) +
    scale_y_discrete(expand = c(0, 0), position = "right") +
    coord_flip() +
    facet_grid(comparisonType ~ .,
        scale = "free_y",
        space = "free_y"
    ) +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.spacing = unit(-0.1, "lines"),
        strip.background = element_blank(),
        axis.line = element_blank(),
        strip.text.y = element_blank()
    )

p_safety_main <- plot_grid(
    p_safety_main_dots +
    theme(plot.margin = margin(
        t = 5.5, b = 5.5, l = 5.5, r = -4, unit = "pt")),
    p_safety_main_numbers +
    theme(plot.margin = margin(
        t = 5.5, b = 5.5, l = 0, r = 5.5, unit = "pt")),
    rel_widths = c(1, 0.65),
    align = "h",
    axis = "t",
    nrow = 1,
    ncol = 2
)
p_safety_main

```

### Safety by somatic T-D evidence

```{r, safety_by_cancer_datasource}
cancer_ds <- c("cancer_gene_census", "intogen", "eva_somatic")
p_safety_by_cancer_datasource <- all_tests %>%
    filter(predictionType %in% c("metareason", "clinical", "stopStatus")) %>%
    filter(comparison %in% cancer_ds) %>%
    filter(comparisonType == "byDatasource") %>%
    mutate(predictionType = fct_relevel(predictionType,
    "clinical", "stopStatus", "metareason")) %>%
    mutate(prediction = fct_reorder(prediction, estimate, median)) %>%
    # mutate(comparison = fct_rev(comparison)) %>%
    ggplot(aes(x = prediction, y = estimate)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
    width = 0) +
    geom_point() +
    scale_y_log10(name = "Odds Ratio (log)") +
    coord_flip() +
    facet_grid(
        predictionType ~ comparison,
        scales = "free_y",
        space = "free_y") +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.title.y = element_blank(),
        # strip.text.y = element_blank(),
        strip.background = element_rect(fill = NA),
        panel.grid.major.y = element_blank()
    )
p_safety_by_cancer_datasource
```

### Safety by target genetic constrain

```{r, safety_by_genetic_constrain, cache = TRUE}
p_safety_by_genetic_constrain <- all_tests %>%
    filter(comparisonType == "geneticConstrain") %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    ggplot(aes(
        x = comparison,
        y = estimate,
        color = prediction)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
    width = 0) +
    geom_point() +
    scale_color_discrete(name = "Study stop reason") +
    scale_y_log10(name = "Odds Ratio (log)") +
    scale_x_discrete(name = "pLOEUF target genetic constrain bin (GnomAD)") +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        legend.position = c(0.95, 0.95),
        legend.justification = c(1, 1),
    )
p_safety_by_genetic_constrain
```

### RNA specificity

- Tissue Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to all other tissues.
- Group Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a group of 2-7 tissues compared to all other tissues, and that are not considered Tissue Enriched.
- Tissue Enhanced: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to the average levels in all other tissues, and that are not considered Tissue Enriched or Group Enriched.

```{r, safety_rna_specificity, cache = TRUE}
p_safety_rna_specificity <- all_tests %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    filter(comparisonType == "rnaSpecificity") %>%
    mutate(comparison = fct_relevel(comparison,
    "Low tissue specificity",
    "Tissue enhanced",
    "Group enriched",
    "Tissue enriched")) %>%
    ggplot(aes(
        x = comparison,
        y = estimate,
        color = prediction)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
    width = 0) +
    geom_point() +
    scale_color_discrete(name = "Study stop reason") +
    scale_y_log10(name = "Odds Ratio (log)") +
    scale_x_discrete(name = "RNA tissue specificity") +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        legend.position = c(0.95, 0.95),
        legend.justification = c(1, 1),
    )
p_safety_rna_specificity
```
